Asimilación de distintas observaciones

### 2021-01-11 analysis/chapters/experimento_20181122.Rmd
Update graficos

fisica <- data.table(ens = 1:60, 
                     fisica = rep(c("KF-YSU", 
                                    "BMJ-YSU",
                                    "GF-YSU",
                                    "KF-MYJ",
                                    "BMJ-MYJ",
                                    "GF-MYJ",
                                    "KF-MYNN2",
                                    "BMJ-MYNN2",
                                    "GF-MYNN2"), length.out = 60)) %>% setDT()

@@

```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots).", fig.height=4}
obsAUT <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
                na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
  .[, c("V2", "V4") := NULL] %>% 
  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))

@@

rbindlist(list(oficiales, no_oficiales, ascat, sondeos, aircraf, satwnd)) %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_mapa() +
  geom_point(alpha = 1, size = 0.5,
             aes(color = source, shape = source)) +
  scale_color_brewer(palette = "Dark2") +
  # khroma::scale_color_vibrant() +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")

rbindlist(list(no_oficiales, oficiales)) %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_mapa() +
  geom_point(alpha = 0.8, size = 1,
             aes(color = source)) +
  scale_color_manual(name =  NULL, values = c("Sfc - Non-official" = "#00695c", 
                                              "Sfc - Official" = "#FD8002"),
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("Non-official (N = 820)","Official (N = 327)"))+
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

@@

rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel == 335 & 
      qc == 0 & iuse == 1] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(trans = mesoda::symlog_trans(),
                        guide = guide_colourbar(barheight = 15, barwidth = 1)) +
  geom_mapa() +
  facet_grid(run~date, labeller = labeller(run = c("diagfiles" = "OmB",
                                                   "diagfiles_ana" = "OmA"))) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_minimal() +

  rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel == 335 & 
      # qc == 0 & 
      iuse == 1] %>% 
  dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
  .[, ":="(tbc = diagfiles - diagfiles_ana,
           run = "AmB")] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(guide = guide_colourbar(barheight = 15, barwidth = 1)) +
  geom_mapa() +
  labs(x = NULL, y = NULL, color = NULL) +
  facet_grid(run~date) +
  theme_minimal() +
  plot_layout(ncol = 2, widths = c(0.4, 0.6))
```

@@

@@

@@

#### FSS for the analysis

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_E[4-7]new.csv")

>>>>

fisica <- data.table(mem = as.character(formatC(1:60, flag = "0", width = 3)),
                     fisica = rep(c("KF-YSU", 
                                    "BMJ-YSU",
                                    "GF-YSU",
                                    "KF-MYJ",
                                    "BMJ-MYJ",
                                    "GF-MYJ",
                                    "KF-MYNN2",
                                    "BMJ-MYNN2",
                                    "GF-MYNN2"), length.out = 60)) %>% setDT()

@@

```{r mapa-obs, fig.cap="Localization of meteorological stations and ships for official prep-BUFR (orange dots) and automatic statios (green dots)."}
obsAUT <- fread(here("analysis", "data", "raw_data", "E5_asim_conv_20181121120000.ensmean"), 
                na.strings = c("0.100E+11", "-0.100E+06", "-99999.90", "-100000.00")) %>% 
  .[, c("V2", "V4") := NULL] %>% 
  setnames(colnames(.), c("var", "stationID", "type", "dhr", "lat", "lon", "pressure", "usage.flag", "flag", "obs", "obs.guess", "obs2", "obs.guess2", "rerr"))

@@

rbindlist(list(no_oficiales, oficiales)) %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_mapa() +
  geom_point(alpha = 0.8, size = 1,
             aes(color = source)) +
  scale_color_manual(name =  NULL, values = c("Sfc - Non-official" = "#00695c", 
                                              "Sfc - Official" = "#FD8002"),
                     breaks = c("Sfc - Non-official", "Sfc - Official"),
                     labels = c("Non-official (N = 820)","Official (N = 327)"),
                     guide = guide_legend(override.aes = list(size = 2)))+
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme_minimal() + 
  theme(legend.position = "bottom") +

  plot_spacer() +

  rbindlist(list(ascat, sondeos, aircraf, satwnd)) %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_mapa() +
  geom_point(alpha = 1,
             aes(color = source, shape = source, size = source)) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4", "grey70"),
                     guide = guide_legend(override.aes = list(size = 2))) +
  scale_size_manual(values = c(0.7, 0.5, 1.5, 0.1)) +
  # scale_color_brewer(palette = "Dark2") +
  # khroma::scale_color_vibrant() +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL, color = NULL, shape = NULL, size = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

  plot_layout(ncol = 3, widths = c(0.47, 0.06, 0.47))
```

@@

rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel == 335 & 
      errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200)] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(trans = mesoda::symlog_trans(),
                        guide = guide_colourbar(barheight = 15, barwidth = 1)) +
  geom_mapa() +
  facet_grid(run~date, labeller = labeller(run = c("diagfiles" = "OmB",
                                                   "diagfiles_ana" = "OmA"))) +
  labs(x = NULL, y = NULL, color = NULL) +
  theme_minimal() +

  rad %>% 
  .[sensor == "iasi_metop-b" & 
      channel == 335 & 
      errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200)] %>% 
  dcast(lon + lat + date ~ run, value.var = c("tbc")) %>% 
  .[, ":="(tbc = diagfiles - diagfiles_ana,
           run = "AmB")] %>% 
  .[, date := factor(date)] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = tbc)) +
  scale_color_divergent(guide = guide_colourbar(barheight = 15, barwidth = 1)) +
  geom_mapa() +
  labs(x = NULL, y = NULL, color = NULL) +
  facet_grid(run~date) +
  theme_minimal() +
  plot_layout(ncol = 2, widths = c(0.4, 0.6))
```
```{r omb-qv-acum}

global_qv_E6 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E6.rds")
global_qv_E7 <- read_rds("~/mesoda/analysis/data/derived_data/omb_acum_qv_E7.rds")

rbind(global_qv_E6, global_qv_E7) %>%
  # .[exp == "E7"] %>% 
  .[, nivel := fcase(bottom_top %in% c(1:8), "1 to 8",
                     bottom_top %in% c(9:13), "9 to 13",
                     bottom_top %in% c(14:20), "14 to 20",
                     bottom_top %in% c(21:38), "21 to 38")] %>%
  .[, acum := cumsum(qv), by = .(bottom_top, exp)] %>%
  # .[bottom_top %in% c(1:18)] %>%
  ggplot(aes(date, acum*1000)) +
  geom_line(aes(color = exp, linetype = exp, group = interaction(bottom_top, exp))) +
  # geom_label(data = ~.x[date %in% sample(unique(date), 38)] %>% group_by(exp, bottom_top) %>% sample_n(size = 1),
   geom_label(data = ~.x[date %in% unique(date)[67]],
             aes(label = bottom_top, color = exp, 
                 group = interaction(bottom_top, exp)),
             size = 3, fill = "white", show.legend = FALSE,
             label.size = NA, label.padding = unit(0.1, "lines")) +
  scale_date() +
  scale_linetype_manual(values = c(2, 1), labels = c(E4 = "CONV", E5 = "AUT",
                                     E6 = "SATWND", E7 = "RAD")) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  facet_wrap(~nivel, scales = "free_y") +
  labs(x = NULL, y = "Accumulated specific humidity (g/Kg)",
       color = NULL, linetype = NULL) +
  theme_minimal()

```{r}
files <- Sys.glob("~/datosmunin/EXP/E7/ANA/2018112*/diagfiles/asim*.ensmean")

files <- files[str_detect(files, "conv", negate = TRUE)]

rad <- read_diag_rad(files, "E7") %>% 
  satinfo[., on = c("sensor", "channel")] %>% 
  .[,  type :=  fifelse(sensor %in% multiespectrales, "multiespectral", "otro")] %>% 
  separate(sensor, into = c("sensor", "plat"), sep = "_")

rad[errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200), 
    .(tbc_mean = mean(tbc),
      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
  .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
  # .[, range(diff, na.rm = TRUE)]
  melt(id.vars = c("sensor", "channel", "date")) %>% 
  ggplot(aes(value)) +
  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
                 alpha= 0.3, position = "identity") +
  geom_density(aes(color = variable, y = stat(density))) +
  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  scale_fill_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
  scale_color_viridis_d(labels = c("OmB bc", "OmB nobc", "Difference")) +
  geom_vline(xintercept = 0, color = "grey40") +
  coord_cartesian(xlim = c(-5, 5)) +
  facet_wrap(~sensor) +
  labs(fill = NULL, x = NULL, color = NULL) +
  theme_minimal()

rad[errinv %between% c(0.0000316227, 1000) & 
      qc == 0 & 
      iuse == 1 &
      peakwt %between% c(0.001, 1200), 
    .(tbc_mean = mean(tbc),
      tbcnob_mean = mean(tbcnob)), by = .(sensor, channel, date)] %>% 
  .[, diff := abs(tbcnob_mean) - abs(tbc_mean)] %>% 
  # .[, range(diff, na.rm = TRUE)]
  melt(id.vars = c("sensor", "channel", "date")) %>% 
  ggplot(aes(value)) +
  geom_histogram(aes(fill = variable, y = stat(density)), binwidth = 0.1,
                 alpha= 0.3, position = "identity") +
  geom_density(aes(color = variable)) +
  # scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  scale_color_viridis_d(labels = c("OmB BC", "OmB NoBC", "Difference")) +
  scale_fill_viridis_d(labels = c("OmB BC", "OmB NoBC", "Difference")) +
  geom_vline(xintercept = 0, color = "grey40") +
  coord_cartesian(xlim = c(-5, 5)) +
  # facet_wrap(~sensor) +
  labs(fill = NULL, x = NULL, color = NULL) +
  theme_minimal()

@@

```{r omb-q-evol, fig.cap="Evolution of the mean OmB for humidity throughout the periodod"}
files <- c(Sys.glob("~/datosmunin/EXP/E5/ANA/*/diagfiles/asim_conv_201811*.mem*"),
           Sys.glob("~/datosmunin/EXP/E7/ANA/*/diagfiles/asim_conv_201811*.mem*"))

  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP/{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.mem{mem}")
  read_diag_conv(f, exp = meta[[1]][["exp"]], member = meta[[1]][["mem"]], variable = c("q")) %>% 
    .[stationID %in% oficiales$stationID] %>% 
    .[]
}) %>% rbindlist() %>% 
  fisica[., on = "mem"] %>% 
  separate(fisica,  into = c("conv", "pbl"))

conv[type %in% c(181, 187) & exp %in% c("E5", "E7") & usage.flag == 1, 
     .(mean_omb = mean(obs.guess, na.rm = TRUE),
       n = .N), 
     by = .(date, exp, pbl, usage.flag)] %>% 
  ggplot(aes(date, mean_omb)) +
  geom_line(aes(color = exp, linetype = pbl)) +
  geom_point(aes(color = exp, size = n/60)) +
  scale_size_area(max_size = 2) +
  scale_linetype_manual(values = c(2, 1, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
  #                                                            "-1" = "rejected"))) +
  scale_date(20181120180000, 20181123120000) +
  labs(y = "Mean of OmB (g/g)", x = NULL, linetype = NULL,
       color = NULL, size = "number\n of obs") +
  theme_minimal()
```

@@

```{r dbz-mean}

@@

```{r dbz-ens}

files <- Sys.glob(paste0(derived_data, "dbz/maxdbz_ana_E[4-7]_0*"))
files <- files[str_detect(files, "20181122000000", negate = TRUE)]

  meta <- unglue(basename(f), "maxdbz_ana_{exp}_{mem}_{date}.nc")
  ReadNetCDF(f, vars = c(lon = "XLONG", lat = "XLAT", "max_dbz")) %>% 
    .[, ":="(exp = meta[[1]][["exp"]],
             mem = meta[[1]][["mem"]],
             date = ymd_hms(meta[[1]][["date"]]))] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] 
}) %>% 
  rbindlist() %>% 
  fisica[., on = "mem"] %>% 
  separate(fisica, into = c("cumulus", "pbl"))

dbz %>% 
  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
  .[date == unique(date)[1]] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
                    proj = norargentina_lambert,
                    breaks = seq(0, 60, 5),) +
  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
                  guide = guide_colorsteps(barheight = 18)) +
  geom_mapa() +
  facet_grid(cumulus~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                   E6 = "SATWND", E7 = "RAD"))) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal() +

  dbz %>% 
  .[, .(max_dbz = mean(max_dbz, na.rm = TRUE)), by = .(cumulus, exp, date, lon, lat, x, y)] %>% 
  .[date == unique(date)[2]] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = max_dbz, fill = stat(level_d)), 
                    proj = norargentina_lambert,
                    breaks = seq(0, 60, 5),) +
  scale_color_dbz(aesthetics = "fill", super = ScaleDiscretised, 
                  guide = guide_colorsteps(barheight = 18)) +
  geom_mapa() +
  facet_grid(cumulus~exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                   E6 = "SATWND", E7 = "RAD"))) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_minimal() +
  plot_layout(ncol = 1)
```
#### FSS for the analysis

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_ana_ens_E[4-7]new.csv")
